### 2019-08-05 _publications/jors/paper.tex
Remake PDF
docs(publication): revise for reviewers' feedback
Implement minor revisions based on reviewer's feedback

1. Wheeler, Daniel; (corresponding author) A\\
2. Keller, Trevor; A\\
3. DeWitt, Stephen J.; B\\
4. Jokisaari, Andrea M.; C\\
5. Schwen, Daniel; C\\
6. Guyer, Jonathan E.; A\\
7. Aagesen, Larry; C\\
8. Heinonen, Olle G.; D\\
9. Tonks, Michael R.; E\\
10. Voorhees, Peter W.; F\\
11. Warren, James A.; G

@@

PFHub is a community effort spearheaded by the Center for Hierarchical
Materials Design at Northwestern University and the National Institute
of Standards and Technology in support of phase-field code
development. The current PFHub deployment~\cite{pfhub} focuses on
improving cross-collaboration between phase-field code developers and
practitioners by providing a standardized set of benchmark
problems~\cite{bm1, bm2} and a workflow for uploading and comparing
benchmark results from different phase-field codes (\emph{e.g.},
FiPy~\cite{fipy}, MMSP~\cite{mmsp}, MOOSE~\cite{moose},
PRISMS-PF~\cite{prisms-pf}).

This paper presents the first deployment of the PFHub framework
including its client-side focused design, how it employs external
services and meta-data about the code base. The paper describes the
relative ease with which other scientific groups might adapt the
framework for their own purposes and deploy using the fully
reproducible Nix environment~\cite{nix}.

@@

The PFHub framework provides a template for other small scientific
communities to host custom content and integrate data from members of
their community. The current deployment provides a facility for
uploading, displaying and comparing results from benchmark problems
supporting phase-field code developers and practitioners. However, the
framework and overall philosophy are broadly transferable to other
communities with some custom configuration and content generation. The
framework uses the Jekyll static website generator~\cite{jekyll} along
with automated front-end processing to eliminate the need for a
CMS~\cite{cmsfree}, which is generally costly to maintain especially
for small scientific communities with limited funding and staffing.
The framework relies on the API, WebSocket and webhook infrastructure
that underpins the modern web and allows external services to have
full-duplex communication between servers and browsers. In particular,
PFHub relies on GitHub's well maintained API and webhook functionality
for external services (such as Travis CI~\cite{travis} and
Staticman~\cite{staticman}).

The workflow for uploading benchmark results relies on third party
tools using the following steps, illustrated in
Figure~\ref{fig:pfhub_website}.
\begin{enumerate}
  \item The users are first required to upload simulation outputs to
    an archival resource (\emph{e.g.}, Figshare\footnote{Certain
      commercial equipment, instruments, or materials (or suppliers,
      or software, ...) are identified in this paper to foster
      understanding. Such identification does not imply recommendation
      or endorsement by the National Institute of Standards and
      Technology, nor does it imply that the materials or equipment
      identified are necessarily the best available for the
      purpose.\label{disclaimer}}~\cite{figshare}) configured with
    permissive cross-origin resource sharing (CORS).
  \item The metadata summarizing each simulation is entered into a
    form on the website, including relevant details such as memory
    usage, run time and links to the data archived in the first step.
  \item Upon submission, the Staticman app~\cite{staticman} submits
    the entered metadata as a pull request against the PFHub
    repository hosted by GitHub. The metadata is stored in a YAML file
    with a unique path in the repository.
  \item Travis CI~\cite{travis} performs linting on the submission and
    then launches a temporary version of the proposed website using
    Surge~\cite{surge}. The PFHub admins can then examine the new
    submission and further changes can be made if necessary.
  \item Once review has been completed to the satisfaction of both the
    uploading scientist and the website maintainers, the pull request
    is merged and served to the World Wide Web using a hosting service
    compatible with GitHub Pages.
\end{enumerate}
A combination of Jekyll templates and CoffeeScript are used to access
and download the data links in the submitted YAML files and then
display the data in interactive plots on the website. The interactive
plots are displayed using the Plotly JavaScript Graphing
Library~\cite{plotly} as it provides a programmable interface and
requires minimal configuration.

The current deployment of PFHub has benchmark specifications
consisting of equations, narrative, plots and code samples, and are
composed in Jupyter Notebooks. The Jupyter Notebooks are included as
static objects in the website after translation into HTML using the
nbconvert tool~\cite{jupyter}. There are currently 7 benchmark
problems each with a number of variations. At the time of writing
there are 108 separate benchmark result uploads~\cite{pfhub} submitted
as pull requests and approved following code review to ensure
compatibility with the website.

@@

@@

@@

\begin{enumerate}
  \item Build the Nix environment from a persistent cache on Travis
    CI, reducing the build time.
  \item Run automated tests on Jupyter Notebooks using
    NBval~\cite{nbval} and Py.test~\cite{pytest}.
  \item Run validation tests on HTML files using
    HTMLProofer~\cite{htmlproofer}.
  \item Lint and test front-end CoffeeScript using appropriate tools.
  \item Display a temporary version of the website using
    Surge~\cite{surge} for visual review.
\end{enumerate}

@@

The PFHub framework can be deployed on any platform supporting Nix,
which includes all contemporary Linux and Mac OS X platforms. Since
the framework is built with Jekyll and automated front-end processing,
it can be deployed on GitHub's Pages infrastructure, which enables
streamlined deployment without the need for any back-end
infrastructure and, thus, is largely platform independent. For
development purposes, a local installation of either Nix (on Linux or
Mac) or Docker (on Linux, Mac or Windows) is required.

@@

Currently, a deployment for a new community has not been attempted
and, thus, the above steps need to be refined and documented.

>>>>

1. Wheeler, Daniel; (corresponding author) A\\
2. Keller, Trevor; A\\
3. DeWitt, Stephen J.; B\\
4. Jokisaari, Andrea M.; C\\
5. Schwen, Daniel; C\\
6. Guyer, Jonathan E.; A\\
7. Aagesen, Larry K.; C\\
8. Heinonen, Olle G.; D\\
9. Tonks, Michael R.; E\\
10. Voorhees, Peter W.; F\\
11. Warren, James A.; G

@@

The phase field method (PFM) describes material interfaces at the
mesoscopic scale between atomic scale models and macroscale
models~\cite{MOELANS2008268}. The PFM is well established and there
are an assortment of code frameworks (\emph{e.g.}, FiPy~\cite{fipy},
MMSP~\cite{mmsp}, MOOSE~\cite{moose}, PRISMS-PF~\cite{prisms-pf})
available for solving the wide variety of phenomena associated with
phase field (e.g. dendritic growth, spinodal decomposition, grain
growth)~\cite{tonks}. However, it is difficult for novice as well as
seasoned phase field practitioners to asses the capability of codes
for different phenomena without extensive prototyping and
groundwork. PFHub aims to provide a low barrier for comparing code
output data using a standard set of metrics.

PFHub is a community effort spearheaded by the Center for Hierarchical
Materials Design at Northwestern University and the National Institute
of Standards and Technology in support of phase-field code
development. The current PFHub deployment~\cite{pfhub} focuses on
improving cross-collaboration between phase-field code developers and
practitioners by providing a standardized set of benchmark
problems~\cite{bm1, bm2} and a workflow for uploading and comparing
benchmark results from different phase-field codes.

Community based scientific efforts often require web services to share
and display data in unique ways between groups and institutions. These
services are difficult to implement due to the groundwork required to
investigate and prototype the many data-sharing and CMS tools
available. The PFHub framework provides a template for other
scientific projects beyond the phase field community. The method
outlined in this paper of using static infrastructure coupled with
small independent third party web services provides a flexible
approach eliminating the initial prototyping and on-going maintenance
required for new infrastructure, while allowing developers to focus on
their unique front-end data views.

This paper presents the first deployment of the PFHub framework
including its client-side focused design, how it employs external
services and metadata about the code base. The paper describes the
relative ease with which other scientific groups might adapt the
framework for their own purposes and deploy using the fully
reproducible Nix environment~\cite{nix}.

@@

The PFHub framework provides a template for other small scientific
communities to host custom content and integrate data from members of
their community. The current deployment (see
Figure~\ref{fig:pfhub_frontpage}) provides a facility for uploading,
displaying and comparing results from benchmark problems supporting
phase-field code developers and practitioners. However, the framework
and overall philosophy are broadly transferable to other communities
with some custom configuration and content generation. The framework
uses the Jekyll static website generator~\cite{jekyll} along with
automated front-end processing to eliminate the need for a
CMS~\cite{cmsfree}, which is generally costly to maintain especially
for small scientific communities with limited funding and staffing.
The framework relies on the API, WebSocket and webhook infrastructure
that underpins the modern web and allows external services to have
full-duplex communication between servers and browsers. In particular,
PFHub relies on GitHub's well maintained API and webhook functionality
for external services (such as Travis CI~\cite{travis} and
Staticman~\cite{staticman}).

The workflow for uploading benchmark results relies on third party
tools using the following steps, illustrated in
Figure~\ref{fig:pfhub_website}.
\begin{enumerate}
  \item The users are first required to upload simulation outputs to
    an archival resource (\emph{e.g.}, Figshare\footnote{Certain
      commercial equipment, instruments, or materials (or suppliers,
      or software, ...) are identified in this paper to foster
      understanding. Such identification does not imply recommendation
      or endorsement by the National Institute of Standards and
      Technology, nor does it imply that the materials or equipment
      identified are necessarily the best available for the
      purpose.\label{disclaimer}}~\cite{figshare}) configured with
    permissive cross-origin resource sharing (CORS).
  \item The metadata summarizing each simulation is entered into a
    form on the website (see Figure~\ref{fig:pfhub_upload}), including
    relevant details such as memory usage, run time and links to the
    data archived in the first step.
  \item Upon submission, the Staticman app~\cite{staticman} submits
    the entered metadata as a pull request against the PFHub
    repository hosted by GitHub. The metadata is stored in a YAML file
    with a unique path in the repository, see~\cite{upload_data} for
    an example.
  \item Travis CI~\cite{travis} performs linting on the submission and
    then launches a temporary version of the proposed website using
    Surge~\cite{surge}. The PFHub admins can then examine the new
    submission and further changes can be made if necessary.
  \item Once review has been completed to the satisfaction of both the
    uploading scientist and the website maintainers, the pull request
    is merged and served to the World Wide Web using a hosting service
    compatible with GitHub Pages.
\end{enumerate}
A combination of Jekyll templates and CoffeeScript are used to access
and download the data links in the submitted YAML files and then
display the data in interactive plots on the website. CoffeeScript is
a higher level language than JavaScript and, thus, more readable
whilst also allowing programming in a more functional manner which
makes data manipulation pipelines both more succinct and easier to
understand. The interactive plots (see Figure~\ref{fig:pfhub_results})
are displayed using the Plotly JavaScript Graphing
Library~\cite{plotly} as it provides a programmable interface and
requires minimal configuration, see~\cite{display_data} for typical
data displayed on the interactive plots.

The current deployment of PFHub has benchmark specifications
consisting of equations, narrative, plots and code samples, and are
composed in Jupyter Notebooks. The Jupyter Notebooks are included as
static objects in the website after translation into HTML using the
nbconvert tool~\cite{jupyter}. There are currently 7 benchmark
problems each with a number of variations. At the time of writing
there are 109 separate benchmark result uploads~\cite{pfhub} submitted
as pull requests and approved following code review to ensure
compatibility with the website.

@@

\begin{figure}
  \includegraphics[width=\textwidth]{pfhub_frontpage.png}
  \caption{Front page of the current version of the deployed PFHub
    website showing links to the wiki and upload pages. The hexbin
    splash provides links to phase field papers and websites. This can
    easily be modified with alternative links and images by editing a
    YAML file.}  \centering
  \label{fig:pfhub_frontpage}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=\textwidth]{pfhub_upload.png}
  \caption{The form used to enter simulation result metadata and then
    upload to the repository via a GitHub pull request allowing the
    data to be checked before pushing to the website.}  \centering
  \label{fig:pfhub_upload}
\end{figure}

\begin{figure}
  \includegraphics[width=\textwidth]{pfhub_results.png}
  \caption{Results comparison page for benchmark 3 (Dendritic
    Growth). The data for each upload is plotted asynchronously and the
    page is not affected if a data link dies.}  \centering
  \label{fig:pfhub_results}
\end{figure}

@@

\begin{enumerate}
  \item Build the Nix environment from a persistent cache on Travis
    CI, reducing the build time.
  \item Run automated tests on Jupyter Notebooks using
    NBval~\cite{nbval} and Py.test~\cite{pytest}.
  \item Run validation tests on HTML files using
    HTMLProofer~\cite{htmlproofer}.
  \item Lint and test front-end CoffeeScript using
    Coffeelint~\cite{coffeelint} and Mocha~\cite{mocha}, respectively.
  \item Display a temporary version of the website using
    Surge~\cite{surge} for visual review.
\end{enumerate}

@@

The PFHub framework can be deployed on any platform supporting Nix,
which includes all contemporary Linux and macOS platforms. Since the
framework is built with Jekyll and automated front-end processing, it
can be deployed on GitHub's Pages infrastructure, which enables
streamlined deployment without the need for any back-end
infrastructure and, thus, is largely platform independent. For
development purposes, a local installation of either Nix (on Linux or
Mac) or Docker (on Linux, Mac or Windows) is required.

@@

Further details on deployment and development of PFHub can be found in
the development guide~\cite{development_guide}.  Currently, a
deployment for a new community has not been attempted and, thus, the
above steps need to be refined and documented.
